---
editor_options: 
  chunk_output_type: console
---
<!--
###############################################################################
###############################################################################
### LICENSE:                                                               ###
### This code is available under a CC BY-NC-SA 4.0 license                 ###
### https://creativecommons.org/licenses/by-nc-sa/4.0/                     ###
###                                                                        ###
### BY  – Credit must be given to the creator                              ###
### NC  – Only noncommercial uses of the work are permitted                ###
### SA  – Adaptations must be shared under the same terms                  ###
###############################################################################
###############################################################################
-->

The exclusion criteria were based on the available literature and analyses of the pilot data. We differentiate between exclusion criteria that were applied to the compressed data (e.g., demographic criteria) and those that were applied to the raw response data (e.g., extremely long response times). When participants were excluded based on compressed data criteria, their data was also removed from the raw data. Figure \@ref(fig:plot-exclusion-pilot) and Figure \@ref(fig:plot-exclusion-full) show the impact of the exclusion criteria when applied to the pilot and full data, respectively.

### Compressed Data {#methods-exclusion-compressed}

The compressed data contains information on summary statistics of the IAT, demographics, and answers to explicit questions (removed during data preprocessing). Participants who did not complete the full IAT were removed from further analyses based on missing $D_{IAT}$ and/or an incomplete `session_status`.

In the new guidelines, @GREENWALD2021 state that the IAT retains its' measurement properties with repeated measurement, although evidence indicates more polarized results at the first IAT measurement (section 2–4). We thus opted to refrain from exclusion based on prior IAT experience.

@GREENWALD2003 showed the statistical benefits of removing participants who made too many fast responses. Participants who responded faster than 300 ms on more than 10% of the trials were excluded from subsequent analyses [Step 5 in Appendix B, @GREENWALD2021]. 

The criteria discussed above focused on aspects inherent to the IAT paradigm, but two additional criteria - based on demographic information - warrant consideration. First, to determine the validity of IAT stimuli, we need to ensure that participants have the highest common ground with regards to concept meaning. We cannot control for this fully but restricting the participant sample to participants who live and grew up in the same country is the best approximation possible for the current data set. We thus excluded participants that did not reside in, nor have citizenship of, the United States of America (USA). We further excluded participants who opted not to provide citizenship and/or residency information.

Second, when participants enter the Project Implicit website, they are asked to provide informed consent. Even though participants may have agreed to the terms and conditions, US States enforce age-limits for the ability to provide consent (ranging from 16 – 18; [https://www.ageofconsent.net/states](https://www.ageofconsent.net/states)). We excluded participants who self-reported to be younger than 18 years old at the time of IAT completion. @GREENWALD2021 further proposed that the validity of exemplars may be derived from examination of the data of "young adult subjects" (p. 7). We defined "young adult subjects" as participants between 18 and 25 years old; excluding all participants who self-reported to be older than 25.

### Raw Data {#methods-exclusion-raw}

The raw data includes trial-by-trial information such as response times and errors. Previously slow and/or fast response times were deleted or recoded. In line with guidelines set forth by Greenwald et al. [-@GREENWALD2003; -@GREENWALD2021] we applied an upper limit excluding trials where the response time is larger than 10,000 ms. We did not recode responses that could be considered as "too fast" (e.g., < 400 ms) because systematic too fast responses may offer relevant information about the variation in average response times. We did, however, exclude responses with a response time of zero milliseconds as these were likely the cause of technical malfunction. Note that the trials are removed, but the participants themselves were not excluded.

Contrary to suggested guidelines, in the present research we excluded participants based on error percentages. @GREENWALD2003 showed that it was unnecessary, but not incorrect, to exclude participants with high error rates (Study 3). We opted to exclude participants who performed below chance (< 50% correct trials across all blocks), because the low success-rates may indicate that participants did not understand task demands and should thus not be considered when validating task content.

```{r plot-exclusion-pilot, include = TRUE, fig.cap = "(ref:plot-exclusion-criteria)"}
combined_exclusion_plots(
  IAT_name = "Gender-Career",
  year_of_interest = 2019
)

knitr::include_graphics(here::here(
  "plots",
  "exclusion_criteria",
  "Gender-Career_2019.png"
))
```

<!-- figure caption -->
(ref:plot-exclusion-criteria) Summary of excluded **pilot** data. (A) The effects of the exclusion criteria on the ***compressed*** data. *Completion*: participants were excluded if they failed to complete the full IAT. *D-score*: for various reasons, among which technical error, IAT D-scores (i.e., $D_{IAT}$) may have been missing. *PCT 300*: participants were excluded if the percentage of responses faster than 300 ms was higher than 10%. *Nationality*: participants were excluded if they did not reside in, nor have citizenship of, the United States of America. *Age*: participants were excluded if they were younger than 18 or older than 25. *Total*: the unique number of participants excluded based on missing data and/or explicit exclusion. (B) The effects of the exclusion criteria in the ***raw*** pilot data. *Latency*: trials were excluded if they exceeded 10,000 ms or were 0 ms. *% Errors*: trials were excluded if the participant answered more than 50% of their trials incorrectly. *Total*: the total number of trials excluded based on the prior criteria. **Note:** the data of participants which were excluded based on the *compressed* data were excluded from the *raw* data *prior* to applying the exclusion criteria. After mutual exclusion, `r pilot$nrow_compressed` participants were eligible for analyses.

```{r plot-exclusion-full, include = TRUE, fig.cap ="(ref:plot-full-exclusion-criteria)"}
# CREATE PLOT
combined_exclusion_plots(
  IAT_name = "FULL_SAMPLE",
  year_of_interest
)

# IN-TEXT Descriptives
stats <-
  get_full_sample_descriptives(year_of_interest)

# INCLUDE plot in manuscript
knitr::include_graphics(
  here::here(
    "plots",
    "exclusion_criteria",
    glue::glue(
      "FULL_SAMPLE_{year_of_interest}.png"
    )
  )
)
```

<!-- figure caption -->
(ref:plot-full-exclusion-criteria) Summary of the excluded **full** data. (A) The effects of the exclusion criteria on the ***compressed*** data. *Completion*: participants were excluded if they failed to complete the full IAT. *D-score*: for various reasons, among which technical error, IAT D-scores (i.e., $D_{IAT}$) may have been missing. *PCT 300*: participants were excluded if the percentage of responses faster than 300 ms was higher than 10%. *Nationality*: participants were excluded if they did not reside in, nor have citizenship of, the United States of America. *Age*: participants were excluded if they were younger than 18 or older than 25. *Total*: the unique number of participants excluded based on missing data and/or explicit exclusion. (B) The effects of the exclusion criteria in the ***raw*** data. *Latency*: trials were excluded if they exceeded 10,000 ms or were 0 ms. *% Errors*: trials were excluded if the participant answered more than 50% of their trials incorrectly. *Total*: the total number of trials excluded based on the prior criteria. **Note:** the data of participants which were excluded based on the *compressed* data were excluded from the *raw* data *prior* to applying the exclusion criteria. After exclusion, `r stats$total_included` participants were included in total (N = `r stats$min_pp_per_IAT` - `r stats$max_pp_per_IAT`, $M$ = `r stats$mean_pp_per_IAT`, $SD$ = `r stats$sd_pp_per_IAT`).
