---
editor_options: 
  chunk_output_type: console
---
<!--
################################################################################
################################################################################
### LISCENSE:                                                                ###
### This code is available under a CC BY-NC-SA 4.0 license                   ###
### https://creativecommons.org/licenses/by-nc-sa/4.0/                       ###
###                                                                          ###
### BY  – Credit must be given to the creator                                ###
### NC  – Only noncommercial uses of the work are permitted                  ###
### SA  – Adaptations must be shared under the same terms                    ###
################################################################################
################################################################################
-->

The exclusion criteria are based on the available literature and analyses of the pilot data. We differentiate between exclusion criteria that can be derived from the compressed data (e.g., demographic criteria) and those that can only be derived from the response data (e.g., extremely long response times). When participants are excluded based on compressed data criteria, their data is also removed from the raw data, and vice versa. Figure \@ref(fig:plot-exclusion-pilot) and Figure \@ref(fig:plot-exclusion-full) show the impact of the exclusion criteria when applied to the pilot and full data, respectively.

### Raw Data {#methods-exclusion-raw}

The raw data includes trial-by-trial information such as response times and errors. Previous research excluded slow and/or recoded fast response times. In line with guidelines set forth by Greenwald et al. [-@GREENWALD2003; -@GREENWALD2021] we will apply an upper limit excluding trials where the response time is larger than 10,000 ms. We will not recode responses that could be considered as "too fast" (e.g., < 400 ms) because systematic too fast responses may offer relevant information about the variation in average response times. We will exclude responses with a response time of zero milliseconds as these are likely the cause of technical malfunction. Note that the trials are removed, but the participants themselves are not excluded.

Contrary to suggested guidelines, in the present research we will exclude participants based on error percentages. @GREENWALD2003 showed that it was unnecessary, but not incorrect, to exclude participants with high error rates (Study 3). We opted to exclude participants who performed below chance (< 50% correct trials across all blocks), because the low success-rates may indicate that participants did not understand task demands and should thus not be considered when validating task content.

```{r, pilot-error-rates, eval = FALSE}
# NOTE: NOT included in the manuscript, only for exploratory checks

# GOAL
# Visualize the error rate distribution of the pilot data that caused us to 
# ... specify an exclusion criteria based on error percentages

pilot_error_rates <-
  # Use the raw data PRIOR to exclusion
  pilot_raw_old %>%
        # Compute per participant / session
        group_by(session_id) %>%
        # Determine percentage error
        summarize("n_trials" = n(),
                  # trial_error = 0 (correct) or 1 (incorrect)
                  "n_errors" = sum(trial_error),
                  # computer percentage
                  "perc_errors" = n_errors / n_trials * 100) %>%
        ungroup() 

# compute the quartiles and IQR 
pilot_error_IQR <- IQR(pilot_error_rates$perc_errors) # Inter Quartile Range
pilot_error_Q3 <- quantile(pilot_error_rates$perc_errors)["75%"] # Upper Quartile
pilot_error_upper <- pilot_error_Q3 + (1.5 * pilot_error_IQR) # Outlier Boundary

# Define outlier type (for plotting)
pilot_error_rates <-
  pilot_error_rates %>%
  mutate("outlier_type" = case_when(perc_errors < pilot_error_upper ~ "Normal",
                                    perc_errors <= 50 ~ "Outlier",
                                    # all other cases (ie > 5-)
                                    TRUE ~ "Below Chance"))

pilot_error_n_excluded <- sum(pilot_error_rates$outlier_type != "Normal")
pilot_error_perc_excluded <- round((pilot_error_n_excluded / nrow(pilot_error_rates)) * 100, 2)

# CREATE PLOT
ggplot(data = pilot_error_rates) +
  # VERTICAL LINES
  geom_vline(xintercept = c(pilot_error_upper, 50),
             col = c("#D55E00", "grey")) + # red, grey
  # DATA POINTS
  geom_jitter(aes(x = perc_errors,
                  y = 1,
                  col = outlier_type),
              height = 0.3,
              alpha = 0.4) + 
  # SUMMARIES
  geom_violin(aes(x = perc_errors,
                   y = 2)) + 
  geom_boxplot(aes(x = perc_errors,
                   y = 2),
               width = 0.3,
               outlier.shape = NA) + 
  # STYLING
  theme_minimal() +
  scale_color_manual(values = c("Normal" = "#009E73", # green
                                "Outlier" = "#D55E00", # red
                                "Below Chance" = "black")) + # grey
  scale_x_continuous(breaks = seq(from = 0,
                                  to = 100,
                                  by = 10),
                     labels = function(x) {paste0(x, "%")},
                     sec.axis = sec_axis(trans = ~.,
                                         breaks = c(round(pilot_error_upper, 2), 50),
                                         labels = function(x) {paste0(x, "%")})) +
  labs(x = "Percentage of Incorrect Trials",
       y = element_blank(),
       col = element_blank()) +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        legend.position = "top")

# Descriptives
perc_outlier <- 
  round(sum(pilot_error_rates$outlier_type == "Outlier") / 
          nrow(pilot_error_rates) * 100, 2)
perc_below_chance <-
   round(sum(pilot_error_rates$outlier_type == "Below Chance") / 
          nrow(pilot_error_rates) * 100, 2)

# FIGURE CAPTION
# (ref:plot-pilot-error-rates) The percentage of incorrect trials per participant in the raw pilot data *before* exclusion. Participants have been defined as outliers (red dots) when their error rates exceeded $Q_3 + (1.5 * IQR)$ (red line; `r round(pilot_error_upper, 2)`%). In addition, a special case of outliers was marked when the participants performed below chance (> 50% incorrect; black dots). In total, `r perc_outlier`% of the participants were classed as outliers and `r perc_below_chance`% as performing below chance.
```

### Compressed Data {#methods-exclusion-compressed}

The compressed data contains information on summary statistics of the IAT, demographics, and answers to explicit questions (removed during data preprocessing). Participants who did not complete the full IAT will be removed from further analyses based on missing $D_{IAT}$ and/or an incomplete `session_status`.

Participants were at times excluded who self-reported to have prior IAT experience. In the new guidelines, @GREENWALD2021 state that the IAT retains its' measurement properties with repeated measurement, although evidence indicates more polarized results at the first IAT measurement (section 2–4). We thus opted to refrain from exclusion based on prior IAT experience.

@GREENWALD2003 showed the statistical benefits of removing participants who made too many fast responses. Participants who responded faster than 300 ms on more than 10% of the trials will be excluded from subsequent analyses [Step 5 in Appendix B, @GREENWALD2021]. 

The criteria discussed above focused on aspects inherent to the IAT paradigm, but two additional criteria - based on demographic information - warrant consideration. First, to determine the validity of IAT stimuli, we need to ensure that participants have the highest common ground with regards to concept meaning. We cannot control for this fully but restricting the participant sample to participants who live and grew up in the same country is the best approximation possible for the current data set. We will thus exclude participants that did not reside in, nor have citizenship of, the United States of America (USA). We will further exclude participants who opted not to provide citizenship and/or residency information.

Second, when participants enter the Project Implicit website, they are asked to provide informed consent. Even though participants may have agreed to the terms and conditions, US States enforce age-limits for the ability to provide consent (ranging from 16 – 18; [https://www.ageofconsent.net/states](https://www.ageofconsent.net/states)). We will exclude participants who self-reported to be younger than 18 years old at the time of IAT completion. @GREENWALD2021 further proposed that the validity of exemplars may be derived from examination of the data of "young adult subjects" (p. 7). We define "young adult subjects" as participants between 18 and 25 years old; excluding all participants who self-reported to be older than 25.

```{r, pilot-age-distribution, eval = FALSE}
# NOTE: NOT included in the manuscript, only for exploratory purposes

# Find participant ages
age_dat <-
  pilot_compressed_old %>%
  select(session_id, age) %>%
  na.omit()

# compute the quartiles and IQR 
pilot_age_IQR <- IQR(age_dat$age) # Inter Quartile Range
pilot_age_Q3 <- quantile(age_dat$age)["75%"] # Upper Quartile
pilot_age_upper <- pilot_age_Q3 + (1.5 * pilot_age_IQR) # Outlier Boundary

# CREATE PLOT
ggplot(data = age_dat) +
  # Outlier line
  geom_vline(xintercept = c(18, pilot_age_upper),
             col = "#D55E00") + # red
  # Individual data points
  geom_jitter(aes(x = age,
                 y = 1,
                 col = ifelse(age > pilot_age_upper | age < 18,
                              "Outlier",
                              "Normal")),
             alpha = 0.4,
             height = 0.5) +
  # Summary
  geom_violin(aes(x = age,
                  y = 2)) +
  geom_boxplot(aes(x = age,
                   y = 2),
               width = 0.3) +
  # STYLING
  theme_minimal() +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        legend.position = "top") + 
  labs(x = "Age",
       y = element_blank(),
       col = element_blank()) +
  scale_color_manual(values = c("Normal" = "#009E73", # green
                                "Outlier" = "#D55E00"# red
                                )) + # grey
  scale_x_continuous(n.breaks = 10,
                     sec.axis = sec_axis(trans = ~.,
                                         breaks = c(16, pilot_age_upper),
                                         labels = c("Age of Consent", "Outlier")))

# FIGURE CAPTION
# (ref:pilot-age-caption) The distribution of participant age in the compressed pilot data *before* exclusion. Participants were defined as outliers (red dots) when they (A) were younger than the age of consent (18), or (B) older than `r pilot_age_upper` ($Q_3 + (1.5 * IQR)$).
```

```{r, plot-exclusion-pilot, fig.cap = "(ref:plot-exclusion-criteria)"}
# See functions/exclusion_plot.R

#### RAW ####
# Create the plot
plot_excl_raw <- 
  plot_exclusion_criteria(exclusion_data = pilot_raw_exclusion,
                          datType = "raw")

#### COMPRESSED ####
# Create the plot
plot_excl_compressed <- 
  plot_exclusion_criteria(exclusion_data = pilot_compressed_exclusion,
                          datType = "compressed")

#### COMBINE ####
# Extract the legend, so it can be plotted only once
plot_legend <- cowplot::get_legend(plot_excl_compressed)
# Combine the two plots
cowplot::plot_grid(plot_legend, # first plot the legend
                   # plots without legends
                   plot_excl_raw + theme(legend.position = "none"), 
                   plot_excl_compressed + theme(legend.position = "none"),
                   # Three rows (1 legend + 2 plots)
                   nrow = 3,
                   # Assign labels, leave legend label empty
                   labels = c("", "A", "B"),
                   # Assign heights of the plots, legend should take up less space
                   rel_heights = c(0.2, 1, 1),
                   # Align the plots left
                   axis = "l")
```

<!-- figure caption -->
(ref:plot-exclusion-criteria) Summary of excluded data. (A) The effects of the exclusion criteria on the ***raw*** pilot data. *Completion*: participants were excluded if they failed to complete the full IAT. *D-score*: for various reasons, among which technical error, IAT D-scores (i.e., $D_{IAT}$) may have been missing. *PCT 300*: participants were excluded if the percentage of responses faster than 300 ms was higher than 10%. *Nationality*: participants were excluded if they did not reside in, nor have citizenship of, the United States of America. *Age*: participants were excluded if they were younger than 18 or older than 25. *Total*: the unique number of participants excluded based on missing data and/or explicit exclusion. (B) The effects of the exclusion criteria in the ***compressed*** pilot data. *Latency*: trials were excluded if they exceeded 10,000 ms or were 0 ms. *% Errors*: trials were excluded if the participant answered more than 50% of their trials incorrectly. *Total*: the total number of trials excluded based on the prior criteria. **Note:** the data of participants which were excluded based on the *raw* data were excluded from the *compressed* data, and vice versa. After mutual exclusion, `r nrow(pilot_compressed)` participants were eligible for analyses.

```{r, plot-exclusion-full, fig.cap = "(ref:plot-exclusion-criteria-compressed)", fig.height = 2.5}
# See functions/exclusion_plot.R

# Load exclusion statistics
full_exclusion_stats <-
  read.csv(here::here("datasets",
                      "processed",
                      "compressed_2020_exclusion_statistics.csv"),
           stringsAsFactors = FALSE) %>%
  # Add all numbers together
  group_by(Criteria, Type) %>%
  summarize("N-total" = sum(N),
            "index" = unique(index)) %>%
  ungroup() %>%
  # rename for plotting
  rename("N" = `N-total`)

# CREATE PLOT
plot_exclusion_criteria(exclusion_data = full_exclusion_stats,
                        datType = "compressed")
```

<!-- figure caption -->
(ref:plot-exclusion-criteria-compressed) The aggregated effects of the exclusion criteria when applied to the ***compressed*** data of `r length(included_IATs)` IATs. Even though exclusions based on the raw data (e.g., percentage of errors) are not included, it is noteworthy that the exclusion percentages are comparable to the pilot analyses of the GC-IAT (2019; Figure \@ref(fig:plot-exclusion-pilot)).  *Completion*: participants were excluded if they failed to complete the full IAT. *D-score*: for various reasons, among which technical error, IAT D-scores (i.e., $D_{IAT}$) may have been missing. *PCT 300*: participants were excluded if the percentage of responses faster than 300 ms was higher than 10%. *Nationality*: participants were excluded if they did not reside in, nor have citizenship of, the United States of America. *Age*: participants were excluded if they were younger than 18 or older than 25. *Total*: the unique number of participants excluded based on missing data and/or explicit exclusion.
