---
editor_options: 
  chunk_output_type: console
---
<!--
################################################################################
################################################################################
### LISCENSE:                                                                ###
### This code is available under a CC BY-NC-SA 4.0 license                   ###
### https://creativecommons.org/licenses/by-nc-sa/4.0/                       ###
###                                                                          ###
### BY  – Credit must be given to the creator                                ###
### NC  – Only noncommercial uses of the work are permitted                  ###
### SA  – Adaptations must be shared under the same terms                    ###
################################################################################
################################################################################
-->

```{r, pilot-results-demographics}
mean_age <- round(mean(pilot_compressed$age), 2)
sd_age <- round(sd(pilot_compressed$age), 2)
age_ci <- round(Rmisc::CI(pilot_compressed$age), 2)
```

We conducted pilot analyses with the data of the GC-IAT to formalize the analyses and reports of the results for individual IATs. The results included here will be generated for each of the `r length(included_IATs)` IATs included in the final analysis. The pilot data included `r nrow(pilot_compressed)` participants who provided `r nrow(pilot_raw)` responses overall and `r nrow(pilot_raw[pilot_raw$block_number %in% c(1, 2), ])` responses in Block 1 and 2. The included participants were on average `r mean_age` years old ($SD$ = `r sd_age`; 95% CI = `r age_ci[1]`, `r age_ci[3]`).  

```{r, pilot-analyses}
#### PREPARE DATA ####
pilot_raw_12 <-
  pilot_raw %>%
  # Greenwald et al., 2020: "The useful data will come from Blocks 1 and 2
  # ... of the standard procedure"
  filter(block_number %in% c(1, 2)) %>%
  # Convert to list to allow for repeated selection of the same participant
  # It is ultimately more efficient when applying bootstrap m out of n resamples
  group_by(session_id) %>%
  group_split()

#### CATEGORIES ####
# For plotting
categories <-
  pilot_raw %>%
  group_by(trial_name) %>%
  summarize("category" = unique(block_name)) %>%
  ungroup()

#### BOOTSTRAPS ####
# see functions/bootstrapped_parameters.R
# NOTE: depending on the n_samples this can take a fairly long time to run!
pilot_resamples <-
  bootstrapped_parameters(raw_dat_list = pilot_raw_12,
                          n_samples = n_samples, 
                          n_subjects_per_sample = n_subjects_per_sample)

#### EXEMPLAR VALIDITY ####
# Reliability of the validity estimates is determined from the percentage of 
# ... samples that has reached the sample conclusions
pilot_exemplar_validity <-
  # see functions/validity_estimates.R
  exemplar_validity_estimates(samples = pilot_resamples,
                              perc_limit = 95) %>%
  # Add the category information for plotting
  left_join(categories,
            by = "trial_name") %>%
  # Restructure for ease of interpretation
  arrange(category, trial_name)
```

```{r, pilot-results-text}
#### RESULTS TEXT ####
# see functions/results_text.R
validity_text <- results_text(exemplar_stats = pilot_exemplar_validity,
                              IAT = "GC-IAT")
```

@GREENWALD2021 propose that valid stimuli are those that are rapidly (RT < 800 ms) and accurately (errors < 10%) categorized (p. 7). Figure \@ref(fig:pilot-results-plot) shows the distribution of average response times and error rates across `r n_samples` samples of `r n_subjects_per_sample` participants of the pilot GC-IAT. `r validity_text$RT` `r validity_text$ERROR` `r validity_text$TOTAL`

<center>

```{r, pilot-results-plot, fig.cap = "(ref:results-cap)", warning=FALSE}
#### PLOTS ####
pilot_plots <- 
  results_plot(exemplar_samples = pilot_resamples,
               exemplar_stats = pilot_exemplar_validity,
               filename = here::here("plots",
                                     "pilot-results.png"))
  
knitr::include_graphics(path = here::here("plots",
                                          "pilot-results.png"))  
       
```

</center>

(ref:results-cap) The validity estimates per stimulus of the Gender-Career IAT (November - December, 2019). **(A)** The distribution of *average response time* across `r n_samples` samples of `r n_subjects_per_sample` participants. Within each sample a stimulus (left y-axis) is judged as valid if the average response time is lower than 800 milliseconds (vertical blue line). A stimulus is classed as reliably valid (green) if 95% or more of the samples resulted in a valid judgment (right y-axis) and reliably invalid (red) if a stimulus is classed as valid in 5% or less of the samples. Stimuli that were valid in 6% to 94% of the samples were classed as unreliable (black). **(B)** The distribution of the *percentage of errors* across `r n_samples` samples of `r n_subjects_per_sample` participants. Within each sample a stimulus (left y-axis) is judged as valid if less than 10 percent of the trials were answered incorrectly (vertical blue line). **(C)** An overview of the validity judgments per exemplar based on average response times (RT) and percentage of errors (ERROR). A stimulus was classed as valid (OVERALL) if both criteria were also reliably valid, but as invalid if either criterion was reliably invalid. If the criteria were both unreliable, a stimulus was also classed as unreliable.
